---
title: "p8105_hw1_sg3891"
author: "Sandya Ganesh"
date: 2021-09-28
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Question 1

The following code produces a data frame comprised of:
a random sample of size 10 from a standard Normal distribution, 
a logical vector indicating whether elements of the sample are greater than 0,
a character vector of length 10,
and a factor vector of length 10, with 3 different factor “levels”

```{r df_problem1}
df_problem1 = tibble(
  vec_numeric = rnorm(10),
  vec_logical = vec_numeric > 0,
  vec_char = c("My", "name", "is", "Sandya", "Ganesh", "and", "I", "love",
               "data", "science"),
  vec_factor = factor(c("high", "medium", "low", "medium", "high", "low",
                        "low", "medium", "high", "high"))
)
```

#### Take the mean of each variable in df_problem1

Taking the mean of the numerical random sample vector (vec_numeric) and 
logical vector (vec_logical) worked. However, taking the mean of the 
character vector (vec_char) and factor vector (vec_fac) did not work, and I 
received a warning message stating that the argument is not numeric or logical.

#### Convert the variables in df_problem 1 to numeric

```{r numeric_convert, results = FALSE, warning = FALSE}
as.numeric(pull(df_problem1, vec_logical))
as.numeric(pull(df_problem1, vec_char))
as.numeric(pull(df_problem1, vec_factor))
```

We are able to convert the logical vector to a numeric vector with 0's and 1's.
We are also able to convert the factor vector to a numeric vector with 1, 2, 3.
However, we are not able to convert the character vector to a numeric vector.
This is because there are no commonalities found in the character type, and it
explains why we cannot get a mean for the character vector. However, a mean
was not produced for the factor vector even though it was successfully converted
to a numeric vector, and this may have been because a mean value would not
have been meaningful.


