---
title: "p8105_hw1_sg3891"
author: Sandya Ganesh
date: 2021-09-28
output: github_document
---

Load the tidyverse library
```{r setup, message = FALSE}
library(tidyverse)
```

## Question 1

The following code chunk produces a data frame comprised of:
a random sample of size 10 from a standard Normal distribution, 
a logical vector indicating whether elements of the sample are greater than 0,
a character vector of length 10,
and a factor vector of length 10, with 3 different factor “levels”

```{r df_problem1}
df_problem1 = tibble(
  vec_numeric = rnorm(10),
  vec_logical = vec_numeric > 0,
  vec_char = c("My", "name", "is", "Sandya", "Ganesh", "and", "I", "love",
               "data", "science"),
  vec_factor = factor(c("high", "medium", "low", "medium", "high", "low",
                        "low", "medium", "high", "high"))
)
```

#### Take the mean of each variable in df_problem1

```{r mean_problem1}
mean(pull(df_problem1, vec_numeric))
mean(pull(df_problem1, vec_logical))
mean(pull(df_problem1, vec_char))
mean(pull(df_problem1, vec_factor))
```
Taking the mean of the numerical random sample vector (vec_numeric) and 
logical vector (vec_logical) worked. However, taking the mean of the 
character vector (vec_char) and factor vector (vec_fac) did not work, and I 
received a warning message stating that the argument is not numeric or logical.

#### Convert the variables in df_problem 1 to numeric

```{r numeric_convert, results = FALSE, warning = FALSE}
as.numeric(pull(df_problem1, vec_logical))
as.numeric(pull(df_problem1, vec_char))
as.numeric(pull(df_problem1, vec_factor))
```

We are able to convert the logical vector to a numeric vector with 0's and 1's,
which explains why we can get a mean for the logical vector. This happens 
because the logical vector is outputting TRUE or FALSE, which is easily 
converted to 0's and 1's. 

We are also able to convert the factor vector to a numeric vector with 
1, 2, and 3, because the software recognizes there are 3 levels to the factor
vector and assigns each level a number. However, this does not explain why a 
mean was not produced for the factor vector. This may be because a mean value 
would not have been a meaningful statistic to describe the 3 levels. 

Finally, we are not able to convert the character vector to a numeric vector, 
and it returns an NA value because there is no underlying numeric integer 
representation for the character vector. This explains why we cannot get a mean 
for the character vector, because it is not possible to convert the character 
string into a numeric needed to calculate the mean. 

## Question 2

The following code loads the penguins dataset and includes a description 
of the penguins dataset, as well as a scatterplot of the data.

```{r problem2}
data("penguins", package = "palmerpenguins")
```

#### Description of the Penguins Dataset

Below we describe the dataset including names / values of important variables, 
the size of the dataset, and the mean flipper length.


The names of the variables in the penguins data set include 
`r colnames(penguins)`. Of these variables, some important variables are the
species of penguins, their sex, their bill length, and flipper length. 
There are `r nlevels(pull(penguins, species))` species of penguins that include 
`r levels(pull(penguins, species))`. There are 
`r sum(pull(penguins, sex) == 'female', na.rm = TRUE)` female and
`r sum(pull(penguins, sex) == 'male', na.rm = TRUE)` male penguins in the data.
The range of values for the penguin bill length is 
(`r  range(pull(penguins, bill_length_mm), na.rm = TRUE)` mm). 

The size of the data set penguins is `r nrow(penguins)` rows and 
`r ncol(penguins)` columns.

The mean flipper length is 
`r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE), digits = 2)` mm.

#### Scatterplot of the Penguins Dataset

```{r problem2_histo, warning = FALSE, message = FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)
       ) + geom_point()
ggsave("penguins.png")
```
